An efficient acyclic contact planner for multiped robots

### 2016-12-08 body/intro.tex
ongoin

To make a legged robot walk on a flat ground, efficient approaches based on the capture point(PRATT) can be used to determine the contact location. Choosing the effector
to create a contact with is also trivial because walk follows a cyclic pattern.
Unfortunately, in the general case planning complex contact interactions is extremely challenging.
At any given time a contact choice must be made between infinitely many possibilities (often a combinatorial discrete choice for the effector, but an infinite choice of locations). Furthermore, the contact manifold is foliated~\citep{simeon-manipulation-04}: a contact choice constrains kinematically and dynamically the possible motions, and there is no analytical way to verify whether this decision brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles. Lastly, the contact manifold has a null measure and is thus impossible to randomly sample because it has a zero measure with respect to the configuration space.

@@

%~ Contact interaction is crucial to maintain equilibrium, though extremely challenging:
%~ \begin{enumerate}
%~ \item verifying that a contact allows the motion to be continuously dynamically feasible is computationally expensive; 
%~ \item the contact manifold has a zero measure (a contact configuration cannot be sampled);
%~ \item the choice of contacts along the motion (e.g. left or right hand) adds a combinatorial issue, and the set of possible contact locations is infinite; 
%~ \item kinematic constraints appear upon contact creation, introducing an infinite foliation of the contact manifold~\citep{simeon-manipulation-04};
%~ \item in theory, maintaining the balance of the robot requires the simultaneous planning of the root trajectory and the contact locations and timing, which increases even more the search space.
%~ \end{enumerate}
Most locomotion systems avoid these issues thanks to simplifying assumptions that reduce the complexity (coplanar contacts, cyclic patterns that are not combinatorial)~\citep{Kajita03a}.
At the DARPA challenge, instead of using the environment for creating contacts that facilitate motion, the chosen strategies to locomotion planning avoided obstacles as much as possible~\citep{atkensondarpa}. 
Previous contributions having demonstrated acyclic contact locomotion have not been used because they are either too computationally expensive~\citep{Bretl:2006:MPM:1124573.1124585} or dynamically inconsistent~\citep{Mordatch:2012:DCB:2185520.2185539}.

As suggested by \cite{Bouyarmane2009}, contrary to observation 5 we believe trajectory and contact planning can be decoupled. We go further and claim that this can be done at a much smaller cost, provided we can formulate a computationally-efficient condition for \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} of a root configuration (i.e. there exists a joint configuration such that the robot is in static equilibrium).
This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
This approximation results from a trade-off between a necessary and a sufficient condition for \glslink{equilibrium feasible}{\textit{equilibrium feasibility}}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.

>>>>

To make a legged robot walk on a flat ground, efficient approaches based on the capture point~\citep{Pratt2006} can be used to determine the contact location. Choosing the effector
to create a contact with is also trivial because walk follows a cyclic pattern.
Unfortunately, in the general case planning complex contact interactions is extremely challenging because the determinism disappears.
At any given time a contact choice must be made between infinitely many possibilities (often a combinatorial discrete choice for the effector, and an infinite choice for the location). Furthermore, the contact manifold is foliated~\citep{simeon-manipulation-04}: a contact choice constrains kinematically and dynamically the possible motions, and there is no analytical way to verify whether this decision brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles. Lastly, the contact manifold has a null measure and is thus impossible to randomly sample.

@@

%~ Contact interaction is crucial to maintain equilibrium, though extremely challenging:
%~ \begin{enumerate}
%~ \item verifying that a contact allows the motion to be continuously dynamically feasible is computationally expensive; 
%~ \item the contact manifold has a zero measure (a contact configuration cannot be sampled);
%~ \item the choice of contacts along the motion (e.g. left or right hand) adds a combinatorial issue, and the set of possible contact locations is infinite; 
%~ \item kinematic constraints appear upon contact creation, introducing an infinite foliation of the contact manifold~\citep{simeon-manipulation-04};
%~ \item in theory, maintaining the balance of the robot requires the simultaneous planning of the root trajectory and the contact locations and timing, which increases even more the search space.
%~ \end{enumerate}
%~ Most locomotion systems avoid these issues thanks to simplifying assumptions that reduce the complexity (coplanar contacts, cyclic patterns that are not combinatorial)~\citep{Kajita03a}.
For this reason at the DARPA challenge, instead of using the environment for creating contacts that facilitate motion, the chosen strategies fell back to walking patterns, and avoided obstacles as much as possible~\citep{atkensondarpa}, resulting in unsafe behaviors. 
Previous contributions having demonstrated acyclic contact locomotion have not been used because they are either too computationally expensive~\citep{Bretl:2006:MPM:1124573.1124585} or dynamically inconsistent~\citep{Mordatch:2012:DCB:2185520.2185539}.

In this paper, we propose a pragmatic approach to tackle the complexity of the acyclic contact planning problem. Rather than planning simultaneously in the configuration space the 
whole-body trajectory and the contact location and timings, we decouple path and contact planning, as suggested by \citeauthor{Bouyarmane2009}.
We first plan a path for the geometric root of the robot in a low dimensional space. We formulate this space as a computationally-efficient approximation
of the \textit{equilibrium feasibile} space, that is the space of root configurations for which there exists a joint configuration such that the robot is in static equilibrium.
We then use a deterministic algorithm 
%~ We introduce a computationally-efficient condition for \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} of a root configuration (i.e. there exists a joint configuration such that the robot is in static equilibrium).
%~ We claim that this can be done at a much smaller cost, provided we can formulate a computationally-efficient condition for \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} of a root configuration (i.e. there exists a joint configuration such that the robot is in static equilibrium).
This paper presents a geometrical approximation of this condition, and a concrete implementation of such a decoupled planner.
This approximation results from a trade-off between a necessary and a sufficient condition for \textit{equilibrium feasibility}, allowing us to find solutions extremely rapidly, while preserving a high success rate in the demonstrated scenarios.
